### 2014-09-03 TACO-paper-v2/TACO-Rebuttal.txt
Add rebuttal file

We thank our reviewers (R1-R4). We appreciate the positive feedback and the
critical comments on our paper. Below we answers the various points arose by the
reviewers.