### 2014-10-20 TACO-paper-v2/TACO-Rebuttal-final.txt
Minor fixes to paper. Add a draft of covering letter.

We thank our reviewers (R1-R4). We appreciate the feedback and the critical
comments on our paper obtained in the revision process, which really led
to a significantly improved version. Below we detail how we answered the minor
points raised by R1 and R4.

R1:
1) "I'm not hung up over tables II and III, but putting them in small graphs
would, I think, still be good for the final version, space permitting."

We agree that it would probably be easier to read plots than tables; however,
we also feel that the best way of representing a table such as Tables II and III
is by means of 6 plots (3 plots per architecture, each plot corresponding to a
specific partial differential equation). Due to lack of space, we prefer to stick
to the tables rather than adding a total of 12 tiny plots, which would really
hinder readability.

2) "Also, your comment about different compilation flags (page 13, line 18) is
still not clear, in my opinion.  It reads like other optimization levels were
worse only for the Xeon Phi. If you could just clarify that you picked the best
optimization levels for each platform, that make your methodology more clear."

By re-thinking about this, we actually agree with you: the sentence was not that
clear. So we rephrased it this way, which hopefully avoid any ambiguity:

<< (...). The icc 13.1 compiler was used. We selected the best optimization levels
for each platform; we used -xAVX for auto-vectorization and -O2 on the Sandy Bridge,
and -O3 on the Xeon Phi. >>

R4:
1) "Ideally, we would like COFFEE to be platform independent. However, in this work,
COFFEE is tightly coupled with the target platform. It generates vector intrinsic and
works closely with the PyOP2 runtime. It would be better if the authors could show
that the approach would also work well on a different platform, such as ARM, SPARC,
PowerPC or MIPS. Nevertheless, this is merely a suggestion, not a requirement,
since we understand how much work is involved in tuning for a different platform."

We emphasize that the only way in which COFFEE is tied with the architecture is by means
of vector intrinsics, which are related to the x86 instruction set, while all of the other
transformations occur at the source code level. Using some abstractions on top of vector
intrinsics would probably fix this; this is something we planned for the near future,
since Firedrake is going to be ported to BLUE GENE architectures. COFFEE is
not tied with the PyOP2 runtime either: it is true that a suitable interface is used in
PyOP2 to interact with COFFEE; however, COFFEE can be used in scenarios other than Firedrake/PyOP2,
as demonstrated by the fact that the code lives in a completely different (public)
repository.

2) "The new title of the paper is now changed to focus on cross-loop optimization.
However, for modern microprocessors, cross-loop optimization usually includes locality
improvement transformations to increase the performance of cache hierarchies. COFFEE
may not need to do that since it focuses on lower order methods where the working set
is already fit in the L1 cache. However, the paper should point out why locality
improvement is not in their scope."

We note that locality improvement really is in the scope of the paper: we optimize for
data locality at the level of registers, which is the level of the memory hierarchy
immediately below the L1 cache. In particular, the outer-product vectorization and
expression splitting aim to optimize for register data locality. However, we agree with
you that it is useful to remark this aspect. Therefore, following your adivse, we have
rewritten a few key sentences. Firstly, in the abstract, we changed the following sentence:

<< The compiler manipulates abstract syntax trees generated from a domain-specific language
by introducing domain-aware optimizations and, eventually, produces C code including vector
SIMD intrinsics. >>

into:

<< The compiler manipulates abstract syntax trees generated from a domain-specific
language by introducing domain-aware optimizations for instruction-level parallelism
and register locality. Eventually, it produces C code including vector SIMD
intrinsics. >>

This should aware the reader from the very beginning that locality at the level of
registers is the target of our optimization strategy. Note that the term "(register)
locality" is also reiterated throughout the paper. 
Secondly, we have modified another key sentence in the Introduction (page 3):

<< (...). With such a kernel structure, we focus on aspects like the minimization of
floating-point operations, register allocation and instruction-level parallelism,
especially in the form of SIMD vectorization.>>

into:

<< (...). With such kernel and working set structures, traditional transformations for
cache locality (e.g. blocking) are not helpful; instead we focus on aspects like the
minimization of floating-point operations, register locality and instruction-level
parallelism, especially vectorization.>>

which we think explicitly targets your concern.

3) "The current trend for GPGPU is to have the GPU sharing the memory with the CPU
(so called SVM – Shared Virtual Memory). This SVM feature has already shown up in
OpenCL 2.0 and Cuda 6.0 standards. With this trend, such claims about the stringent
memory requirements of GPU will no longer to be true."

Even though SVM is going to be part of the new GPU programming models (OpenCL, CUDA),
that does not mean that "the stringent memory requirements will no longer be true". SVM
is just an abstraction that allows programmers to avoid explictly handling memory management;
however, behind the scenes, the continuous transfers between host and device will
still impair the performance. We therefore prefer to be really coutious about this matter.

4) "<<There has been some research on adapting local assembly to GPUs (which we mention
in Section 7), although it differs from ours in several ways, including: (i) not relying
on automated code generation from a domain-specific language (explained next),
(ii) testing only very low order methods, (iii) not optimizing for cross-loop arithmetic
intensity (the goal is rather effective multi-thread parallelization). >>

Your work focuses on low order methods, yet you claim that similar work on GPU is different
from yours since they work on “very low” order methods. What are the differences between “low”
and “very low”? From the description of this paper, it seems that they are very close. Some
of the low order methods have trip count s as low as 3, so how much lower can “very low” go?"

We definitely agree our description was unclear. Therefore, we rewrote that part of
the paper as follows:

<<There has been some research on adapting local assembly to GPUs, which we men-
tion in Section 7, although it differs from ours in several ways, including: (i) not op-
timizing for cross-loop arithmetic intensity (the goal is rather effective multi-thread
parallelization), and (ii) not relying on automated code generation from a domain-specific
language (explained next); moreover, they have done testing only on order 1 methods,
while we focus on the order range 1–4. >>

This way not only we clarify the previous ambiguity (by explicitly stating they only test
order 1 methods), also do we emphasize the importance of points (i) and (ii).

5) "Section 6, page 20, line 41
rather then -->  rather than"

Fixed

6) "The treatment of expression split is still too simple. Although real implementations
favor simplicity, a technical paper does not."

For this aspect, we have added a few sentences to Section 3.4, where expression
splitting is described:

<< (...) Nevertheless the performance gain from improved register reuse can still be greater
if suitable heuristics are used. Our approach consists of traversing the expression tree
and recursively splitting it into multiple sub-expressions if the number of innermost-
loop independent variables is larger than a certain threshold. This is elaborated in the
next section, and validated against empirical search in Section 5.2.6. >>

This does not make the treatment less simplistic, but helps understanding the heuristics
we have used (by also adding a forward pointer to the upcoming section, in which COFFEE's
cost model is detailed) and by stating that this heuristics was "validated against empirical
search". The performance evaluation shows that this is indeed a good heuristics, but as
you correctly point out, there is space for further and systematic improvement. That is why
we also added a sentence to the Conclusion section, in which we mention planned further work:

<< (...). As suggested by the performance evaluation, an analysis of more sophisticated
approaches for expression splitting also looks promising.>>

With the expression splitting transformation, our intention was to show that by generalizing
the well-known loop fission transformation, and by using simple intuitions, register locality
can notably be improved. Probably, a formal study would require a complete new paper, because
the complexity of the problem grows considerably if we consider generic expressions embedded
in generic loop nests. We believe that the ideas provided in this paper represent a starting
point for an in-depth, sufficiently abstract analysis of the topic.